
\subsection{Subspaces}

\subsubsection*{}

\begin{definition}
		A subset \(W\) of a vector space \(V\) over a field \(F\) is called a \textit{subspace}\index{subspace} of \(V\) if \(W\) is a vector space over \(F\) with the operations of addition and scalar multiplication defined on \(V\).

\end{definition}


\begin{theorem}[Subspace of a Vector Space]
		Let \(V\) be a vector space and \(W \subseteq V\). Then \(W\) is a subspace of \(V\) if and only if the following conditions hold:

\[
0 \in V
\]

\[
\forall x, y \in W, \; x + y \in W
\]

\[
\forall c \in F, x \in W, \; cx \in W
\]

\end{theorem}


\begin{example}
		The set \(W\) of all symmetric matrices (matrices such that \(A^{t} = A\)) is a subspace of \(M_{m\times n}(F)\). Since the zero matrix has all zero entries, clearly the zero matrix is in \(W\). If \(A,B\in W\), then \((A+B)^{t} = A^{t}+B^{t} = A+B\) so \(W\) is closed under addition. If \(A\in W\) and \(c\in F\), then \(cA^{t} = c(A^{t}) = c(A) = cA\) so \(W\) is closed under scalar multiplication.

\end{example}


\begin{example}
		The set \(W\) of diagonal matrices is a subspace of \(M_{n \times n}(F)\). Clearly, the zero matrix is diagonal so it is in \(W\). Further, if \(A,B\in W\) then \((A+B)_{ij} = A_{ij}+B_{ij}\) and when \(i\not = j\), \(A_{ij} = B_{ij} = 0\). So \(A_{ij}+B_{ij}\) is diagonal and in \(W\). The same argument applies to closure under scalar multiplication.

\end{example}


\begin{theorem}[Intersection of Subspaces]
		Any intersection of subspaces of a vector space \(V\) is a subspace.

\begin{proof}
			Since every subspace contains the zero vector, the intersection of subspaces contains the zero vector. The closure of addition and multiplication follow from the definition of closure.

\end{proof}

\end{theorem}

\subsubsection*{Problems}


\begin{problem}
		Prove that \((A^{t})^{t} = A\) for all \(A \in M_{m\times n}(F)\)

\end{problem}


\begin{solution}
		Let \(A \in M_{m\times n}(F)\). Then by the definition of transpose,

\[
(A^{t})^{t} = (A^{t})_{ij}^{t} = (A_{ji})^{t} = A_{ij} = A
\]

\end{solution}


\begin{problem}
		Prove that diagonal matrices are symmetric matrices

\end{problem}


\begin{solution}
		Let \(A\) be a diagonal matrix. Then where \(i \not = j\), \(A_{ij} = 0\). Additionally, for all \(i = j\), we have \(A_{ij} = A_{ji}\). Then

\[
A^{t} = (A^{t})_{ij} = A_{ji} = A_{ij} = A
\]
		and \(A\) is symmetric by definition.

\end{solution}


\begin{problem}
		Prove that the set \(W_{1} = \{(a_{1}, a_{2},\dots, a_{n} \in F^{n}|a_{1} + a_{2} +\dots+ a_{n} = 0\}\) is a subspace and the space \(W_{2} = \{(a_{1}, a_{2},\dots, a_{n} \in F^{n}|a_{1} + a_{2} +\dots+ a_{n} = 1\}\) is not.

\end{problem}


\begin{solution}
		Let \(W_{1}\) be defined as above. Since \(a_{1} + a_{2} +\dots+ a_{n} = 0\), the zero vector is in \(W_{1}\). Let \(\{a_{n}\}, \{b_{n}\} \in W_{1}\). Then \(\{a_{n}\} + \{b_{n}\} = (a_{1} + a_{2} +\dots+ a_{n}) + (b_{1} + b_{2} +\dots+ b_{n}) = 0 + 0 = 0 \in W_{1}\). For \(c\in F\), \(c\{a_{n}\} = c(a_{1} + a_{2} +\dots+a_{n}) = c(0) = 0 \in W_{1}\). So \(W_{1}\) is a subspace. Define \(W_{2}\) as above, the zero vector is not in \(W_{2}\) so it is not a subspace.

\end{solution}